Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-10-30 papers/thesis-kirstie/thesis.tex
edits

The goal of this project is to generate temperature-density and 
pressure-temperature phase diagrams for a Weeks-Chandler-Anderson (WCA) 
model fluid using a computer program. 
To do this we need to find the Helmholtz free energies that the fluid
would have in a liquid state, and in a crystalline solid state 
at a given temperature and density. 
%Once the Helmholtz free energy for both the crystal and liquid states of 
%the fluid are found for a given temperature and density, they can be 
%compared to find the resulting state of the fluid. 
%
%save:
%Outside the region of 
%coexistence which exists between the liquid and crystal states, and which 
%consists of various amounts of liquid and crystal existing together,
%the fluid will be completely in a crystalline state if the 
%crystal free energy is lower than the liquid free energy, and it will be 
%completely in a liquid state if the liquid free energy is lower than 
%the crystal free energy. Maxwell's Double Tangent Construction is used 
%to determine the actual number densities of the crystal and liquid states
%at the boundaries of the coexistance region for a given temperature from 
%which the phase diagrams can be constructed.
%end save
%
%In general, if at some number density and temperature the crystal free energy 
%is lower than the homogeneous free energy of the liquid, then the fluid will 
%be in the crystalline state. If the homogeneous liquid free energy is lower 
%than the crystal free energy, then the fluid will be in the liquid state. 
%
%To find the equilibrium Helmholtz free energies we use classical Density 
%Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT).
%%the corresponding equilibrium number density profile. 
%In particular, 
In general, the Helmholtz free energy, 
which consists of an ideal part and an excess part ,
can be written as a functional of 
the number density profile $n(\vec{r})$ which describes the way the average number 
density varies spatially. 
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
%

In this paper, classical Density Functional Theory (cDFT)
will be used to find the equilibrium Helmholtz free energy, and 
Soft Fundamental Measure Theory (SFMT) will be used to form the functional for the excess Helmholtz 
free energy. 

@@

According to classical Density Functional Theory (cDFT) the equilbrium 

@@

Soft Fundamental Measure Theory (SFMT) makes the task of forming an expression

@@

The minimization can be illustrated by considering that according to the 
second law of thermodynamics, a system that is not in equilibrium will 
exprience spontanteous processes that increase the total entropy of a 
system and its surroundings until equilibrium is reached, 
$\Delta{S}_{System + Surroundings} \geq 0$. Here ``surroundings" means 
the entire universe minus the system. This, together with the conservation 
of energy (the first law of thermodynamics) which says that energy lost 
(or gained) by the system is gained (or lost) by the surroundings 
$\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that 
changes in the free energy will be negative for spontaneous processes 
when the natural variables of the free energy are held fixed. 
%And so the free energy tends toward a minimum. 
%How will a system at a fixed temperature, volume, and number of 
%particles respond to a small offset from equilibrium?

Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
\begin{equation}
  df_{N-System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}
\end{equation}
and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that 
$dT=0$ gives
\begin{equation}
  f_{N-System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}
\end{equation}
where conservation of energy was also applied. Fixing the volume and the 
number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, 
results in
\begin{equation}
  df_{N-System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
\end{equation}
which, for fixed N, is the same as
\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} 
Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number 
of particles are fixed, the change in the free energy of the system will 
always be negative, and thus the free energy tends toward a 
minimum~\cite{schroeder}. This illustrates the tendancy toward minimization 
for a system near equilibrium. A general proof of minimization 
can be found in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.

@@

The equations developed so far have been for a hard sphere fluid. All of 
the weight function equations, as well as the equation for the deriviative 
of the mayer function, can be expressed in terms of $w_{2}(r)$ which consists 
of a delta function characteristic of hard spheres. Soft Fundamental Measure 
Theory (SFMT) extends the Fundamental Measure Theory discussed so far to 
fluids with soft potenials by keeping the relationships between the weight 
functions, and that the weight functions satisfy the same, but replacing 
$w_{2}(r)$ with a weight function appropriate for a soft sphere potential. 
A new set of weight functions can then be derived for a system of soft spheres.

@@

\section{Overview}
The goal of this project is to generate phase diagrams for a 
Weeks-Chandler-Anderson model fluid using a computer program. 
To do this we find the Helmholtz free energies that the fluid
would have at a given temperature and density in a  
liquid state, and in a crystalline solid state. 
%Once the Helmholtz free energy for both the crystal and liquid states of 
%the fluid are found for a given temperature and density, they can be 
%compared to find the resulting state of the fluid. 
Outside the region of 
coexistence which exists between the liquid and crystal states, 
the fluid will be completely in a crystalline state if the 
crystal free energy is lower than the liquid free energy, and it will be 
completely in a liquid state if the liquid free energy is lower than 
the crystal free energy. Maxwell's Double Tangent Construction is used 
to determine the actual number densities of the crystal and liquid states
at the boundaries of the coexistance region for a given temperature, 
and construct phase diagrams.

%In general, if at some number density and temperature the crystal free energy 
%is lower than the homogeneous free energy of the liquid, then the fluid will 
%be in the crystalline state. If the homogeneous liquid free energy is lower 
%than the crystal free energy, then the fluid will be in the liquid state.

%To find the equilibrium Helmholtz free energies we use classical Density 
%Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT).
%%the corresponding equilibrium number density profile. 
%In particular,

In general, the Helmholtz free energy can be written as a functional of 
the number density profile $n(\vec{r})$ which describes the way the average number 
density varies spatially. 
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
Classical Density Functional Theory (cDFT) is used to find the equilbrium 
Helmholtz free energy of the system at a given temperature and density
by varying the number density profile $n(\vec{r})$ until the free energy 
is minimized. The minimum free energy, and its corresponding number 
density profile are then the equilibrium values for that temperature 
and number density.

The Helmholtz free energy consists of two parts: the ideal Helmholtz free energy
and the excess Helmholtz free energy.
The ideal Helmholtz free energy is the free energy the fluid would have in
the absence of any interactomic potential energy $V(r)$ between the spheres, 
or atoms that make up the fluid. The excess Helmholtz free energy is that 
additional energy which results from the presence of an interatmoic potential 
energy $V(r)$ between the spheres.
%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
%For a spatially varying number density, as would be expected for a crystal, 
%the free energy is a functional of the number density profile $n(\vec{r})$  
%The Helmholtz free energy excess free energy is then added to the ideal free energy calculated 
%analytically to find the resulting free energy for the system. 
Soft Fundamental Measure Theory (SFMT) is used to form the expression 
for the excess Helmholtz free energy without having to take explicitly 
into account the interaction each sphere, or atom, has with every other 
sphere.

 ******

Although the entire crystal lattice can be constructed with FCC cells, 
FCC cells are not the smallest cells with which the lattice can be built. 
The shaded parallepiped shown within the FCC cell in Figure \ref{fig:primitivecell} 
is the smallest structure that, when repeated, re-creates the entire crystal 
lattice. This parallepiped primitive cell takes up one-fourth the volume 
of an FCC cell, and contains the equivalent of one atom.
It includes portions of eight atoms each located on a lattice site convenietly 
reachable by moving along one of three different lattice vectors to step 
from one atom to the next. The three lattice vectors will be useful in 
creating the lattice structure computationally, and will also make the 
program run faster.

@@

It may be that not all lattice sites are occupied. Vacancies can be taken 
into account while keeping %the volume of the bold box, and thus 
the overall number density the same. This is done by 
making the cells smaller. A large collection of smaller cells creates a 
new number density profile shown in Figure \ref{fig:Ensemble_Smallcells} 
with eight smaller Gaussians of one-eigth the height. 
Instead of thinking of an atom in one of the eight small cells, the same 
Gaussian distribution can be obtained from thinking of one-eigth of an 
atom in each small cell. This makes it convenient to describe the new 
number density in terms of the fraction of vacancies, $f_v$:

\begin{displaymath}{ n = (1-f_v){n_{no.vacancies}}}\end{displaymath} 
%\begin{align}
  %n(\vec{r})=(1-f_v)\frac{\exp^{-\frac{{\vec{r}}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}
%\end{align}

@@

By varying the width of the Gaussians and the fraction of vacancies, new 
number density profiles can be generated as shown in Figure~\ref{fig:Ensemble_vary}.
Once again, it is important to note that the number density is the same 
for all these number density profiles $n(\vec{r})$. 
%There is still a total of one atom per ``bold box" volume. 
%This prodcedure is used only to find the crystal free energy as the number density 
%profile for a liquid is homogeneous, as shown in Figure \ref{fig:homogen_denisty}.  
A homogeneous number density profile is shown in Figure \ref{fig:homogen_denisty}.

@@

@@

\color{red}
Figure \ref{fig:weight_functions} shows a 1 dimensional view of the weight functions 
$w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance along the z-axis. In three 
dimensions, the weighting functions $w_0$, $w_1$, $w_2$ are spherical shells 
centered at $\vec{r}$. As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, 
$w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, a width determined 
by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving 
the shell. The weighting function $w_3$, shown in Figure \ref{fig:W0andW3}b is a solid 
sphere up to a radius of ?FIX? after which it tapers off over a distance of ?FIX?.

The parameters $\alpha$ and $\Xi$ are temperature dependent, as shown earlier in 
Figure \ref{fig:alphaXivsT}. $\Xi$ is related to the width of the functions which 
become wider in general as the temperature increases. The radial distance to the 
peak decreases as the temperature increases, that is, the shells shrink, as does 
the sphere. The functions can be treated as neglible outside of a distance of 
about $3(\frac{\Xi}{\sqrt{2}}) + \frac{\alpha}{2}$ CHECK!! from the center of 
the weight function.

 \begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{weight_functions.pdf}
    \caption{The weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
    z-axis for a reduced temperature of 2.}
    \label{fig:weight_functions}
  \end{figure}

 \begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.4\linewidth}
      \includegraphics[height=5cm]{W2.png}
      % \includegraphics[width=5cm]{figs/w2-ring}
      \caption{A crossectional view of $w_2$}
    \end{subfigure} 
    \begin{subfigure}[b]{0.4\linewidth}
      \includegraphics[height=5cm]{W3.png}
      \caption{A crossectional view of $w_3$}
    \end{subfigure} 
    \caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
    spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
    have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
    \label{fig:W0andW3}
    \end{figure} 
\color{black}

\section{Computing the Liquid and Crystal Helmholtz free energies}

>>>>

The goal of this project is to generate computationally temperature-density 
and pressure-temperature phase diagrams for a Weeks-Chandler-Anderson (WCA) 
model fluid using classical Density Functional Theory (cDFT). 
To do this we need to find the Helmholtz free energies that the fluid
would have in a liquid state, and in a crystalline solid state 
at a given temperature and density. 
%Once the Helmholtz free energy for both the crystal and liquid states of 
%the fluid are found for a given temperature and density, they can be 
%compared to find the resulting state of the fluid. 
%
%save:
%Outside the region of 
%coexistence which exists between the liquid and crystal states, and which 
%consists of various amounts of liquid and crystal existing together,
%the fluid will be completely in a crystalline state if the 
%crystal free energy is lower than the liquid free energy, and it will be 
%completely in a liquid state if the liquid free energy is lower than 
%the crystal free energy. Maxwell's Double Tangent Construction is used 
%to determine the actual number densities of the crystal and liquid states
%at the boundaries of the coexistance region for a given temperature from 
%which the phase diagrams can be constructed.
%end save
%
%In general, if at some number density and temperature the crystal free energy 
%is lower than the homogeneous free energy of the liquid, then the fluid will 
%be in the crystalline state. If the homogeneous liquid free energy is lower 
%than the crystal free energy, then the fluid will be in the liquid state. 
%
%To find the equilibrium Helmholtz free energies we use classical Density 
%Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT).
%%the corresponding equilibrium number density profile. 
%In particular, 
In general, the Helmholtz free energy, 
which consists of an ideal part and an excess part ,
can be written as a functional of 
the number density profile $n(\vec{r})$ which describes the way the average number 
density varies spatially. 
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
%

Classical Density Functional Theory is used to find the equilibrium 
Helmholtz free energy, and Soft Fundamental Measure Theory (SFMT) will 
be used to form the functional for the excess Helmholtz free energy. 

@@

According to classical Density Functional Theory the equilbrium 

@@

SFMT makes the task of forming an expression

@@

The minimization can be illustrated by considering that according to the 
second law of thermodynamics, a system that is not in equilibrium will 
exprience spontanteous processes that increase the total entropy of a 
system and its surroundings until equilibrium is reached, 
$\Delta{S}_{System + Surroundings} \geq 0$. Here ``surroundings" means 
the entire universe minus the system. This, together with the conservation 
of energy (the first law of thermodynamics) which says that energy lost 
(or gained) by the system is gained (or lost) by the surroundings 
$\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that 
changes in the free energy will be negative for spontaneous processes 
when the natural variables of the free energy are held fixed~\cite{schroeder}. 
%And so the free energy tends toward a minimum. 
%How will a system at a fixed temperature, volume, and number of 
%particles respond to a small offset from equilibrium? 
%
%Save for me:
%Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
%\begin{equation}
  %df_{N-System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}
%\end{equation}
%and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that 
%$dT=0$ gives
%\begin{equation}
  %f_{N-System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}
%\end{equation}
%where conservation of energy was also applied. Fixing the volume and the 
%number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, 
%results in
%\begin{equation}
  %df_{N-System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
%\end{equation}
%which, for fixed N, is the same as
%\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} 
%Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number 
%of particles are fixed, the change in the free energy of the system will 
%always be negative, and thus the free energy tends toward a 
%minimum~\cite{schroeder}. This illustrates the tendancy toward minimization 
%for a system near equilibrium. 
%End save for me.
A general proof of minimization 
can be found in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.

@@

The equations developed so far have been for a hard sphere fluid. All of 
the weight function equations, as well as the equation for the deriviative 
of the mayer function, can be expressed in terms of $w_{2}(r)$ which consists 
of a delta function characteristic of hard spheres. Soft Fundamental Measure 
Theory extends the Fundamental Measure Theory discussed so far to 
fluids with soft potenials by keeping the relationships between the weight 
functions, and that the weight functions satisfy the same, but replacing 
$w_{2}(r)$ with a weight function appropriate for a soft sphere potential. 
A new set of weight functions can then be derived for a system of soft spheres.

@@

\section{Overview}
%
%The goal of this project is to generate phase diagrams for a 
%Weeks-Chandler-Anderson model fluid using a computer program. 
%To do this we find the Helmholtz free energies that the fluid
%would have at a given temperature and density in a  
%liquid state, and in a crystalline solid state. 
%%Once the Helmholtz free energy for both the crystal and liquid states of 
%%the fluid are found for a given temperature and density, they can be 
%%compared to find the resulting state of the fluid. 
%Outside the region of 
%coexistence which exists between the liquid and crystal states, 
%the fluid will be completely in a crystalline state if the 
%crystal free energy is lower than the liquid free energy, and it will be 
%completely in a liquid state if the liquid free energy is lower than 
%the crystal free energy. Maxwell's Double Tangent Construction is used 
%to determine the actual number densities of the crystal and liquid states
%at the boundaries of the coexistance region for a given temperature, 
%and construct phase diagrams.

%%In general, if at some number density and temperature the crystal free energy 
%%is lower than the homogeneous free energy of the liquid, then the fluid will 
%%be in the crystalline state. If the homogeneous liquid free energy is lower 
%%than the crystal free energy, then the fluid will be in the liquid state. 
%
%%To find the equilibrium Helmholtz free energies we use classical Density 
%%Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT).
%%%the corresponding equilibrium number density profile. 
%%In particular, 
%
%In general, the Helmholtz free energy can be written as a functional of 
%the number density profile $n(\vec{r})$ which describes the way the average number 
%density varies spatially. 
%\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
%Classical Density Functional Theory is used to find the equilbrium 
%Helmholtz free energy of the system at a given temperature and density
%by varying the number density profile $n(\vec{r})$ until the free energy 
%is minimized. The minimum free energy, and its corresponding number 
%density profile are then the equilibrium values for that temperature 
%and number density. 
%
%The Helmholtz free energy consists of two parts: the ideal Helmholtz free energy
%and the excess Helmholtz free energy.
%The ideal Helmholtz free energy is the free energy the fluid would have in
%the absence of any interactomic potential energy $V(r)$ between the spheres, 
%or atoms that make up the fluid. The excess Helmholtz free energy is that 
%additional energy which results from the presence of an interatmoic potential 
%energy $V(r)$ between the spheres.
%%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
%%For a spatially varying number density, as would be expected for a crystal, 
%%the free energy is a functional of the number density profile $n(\vec{r})$  
%%The Helmholtz free energy excess free energy is then added to the ideal free energy calculated 
%%analytically to find the resulting free energy for the system. 
%Soft Fundamental Measure Theory is used to form the expression 
%for the excess Helmholtz free energy without having to take explicitly 
%into account the interaction each sphere, or atom, has with every other 
%sphere.

Our goal is to generate computationally temperature-density and pressure-
temperature phase diagrams for a Weeks-Chandler-Anderson model fluid using 
classical Density Functional Theory. 
A temperature-density phase diagram for a fluid that does not
have attraction between the particles that make up the fluid will show
two discinct regions - one at lower densities, and one at higher densities,
seperated by a coexistence region in which there is found a mixture of the 
fluid at some fixed lower density, and at some fixed higher density. An 
abrupt change from some maximum lower density to some minimum higher density 
occurs for certian combinations of pressure and temperature which form
a line on a pressure-temperature phase diagram. This line distinguishes 
the region in which the fluid exhibits lower densities from the region 
in which the fluid exhbits higher densities. These two regions are known
as the liquid state and the solid state of the fluid.

It is known that a fluid with a soft potential like the WCA fluid will 
form a Face Centered Cubic (FCC) crystal in its solid state~\cite{Hansen}. 
The transition between the liquid 
and solid states is abruptly inhomogeneous, and we study the behavior 
of the fluid by treating the liquid as homogeneous, and by constructing 
a crystal lattice in real space for the solid. Although the entire 
crystal lattice can be constructed with FCC cells, 
FCC cells are not the smallest cells with which the lattice can be built. 
The shaded parallepiped shown within the FCC cell in Figure \ref{fig:primitivecell} 
is the smallest structure that, when repeated, re-creates the entire crystal 
lattice. This parallepiped primitive cell takes up one-fourth the volume 
of an FCC cell, and contains the equivalent of one atom.
It includes portions of eight atoms each located on a lattice site convenietly 
reachable by moving along one of three different lattice vectors to step 
from one atom to the next. The three lattice vectors will be useful in 
creating the lattice structure computationally, and will also make the 
program run faster.

@@

It may be that not all lattice sites are occupied. Vacancies can be taken 
into account while keeping %the volume of the bold box, and thus 
the overall number density the same. This is done by 
making the cells smaller. A large collection of smaller cells creates a 
new number density profile shown in Figure \ref{fig:Ensemble_Smallcells} 
with eight smaller Gaussians of one-eigth the height. 
Instead of thinking of an atom in one of the eight small cells, the same 
Gaussian distribution can be obtained from thinking of one-eigth of an 
atom in each small cell. This makes it convenient to describe the new 
number density in terms of the fraction of vacancies, $f_v$: 
\begin{displaymath}{ n = (1-f_v){n_{no.vacancies}}}\end{displaymath} 
%\begin{align}
  %n(\vec{r})=(1-f_v)\frac{\exp^{-\frac{{\vec{r}}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}
%\end{align}

@@

By varying the width of the Gaussians and the fraction of vacancies, various
number density profiles can be generated as shown in Figure~\ref{fig:Ensemble_vary}.
Once again, it is important to note that the number density is the same 
for all these number density profiles $n(\vec{r})$. 
%There is still a total of one atom per ``bold box" volume. 
%This prodcedure is used only to find the crystal free energy as the number density 
%profile for a liquid is homogeneous, as shown in Figure \ref{fig:homogen_denisty}.  
A homogeneous number density profile is used for a liquid is shown in Figure \ref{fig:homogen_denisty}.

@@

\section{Computing the Liquid and Crystal Helmholtz free energies}
%-----------Comuting the Liquid and Crystal free energies ------

The liquid ideal Helmholtz free energy is computed analytically 
using Equation~\ref{eq:Fideal} 
for a given homogeneous reduced density $n^*$, and reduced temperature $t^*$. 
%from the equation 
%\begin{displaymath}{\frac{F_{Hideal}}{V}=k_BTn(\ln(\frac{n*2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)-DELETE-EQN!}\end{displaymath} 
%\begin{align}
  %\frac{F_{Hideal}}{V}=k_BTn(\ln(n\Lambda^3)-1.0)
%\end{align}
%The quantity $\Lambda$ is the DeBroglie wavelength given by
%\begin{displaymath}{\Lambda =\sqrt{\frac{2\pi\hbar^2}{k_BTm}}}\end{displaymath}  
%where in our program we have set
%\begin{displaymath}\left(\sqrt{\frac{2\hbar^2}{m}}\right)^3=4.752731840013864\text{x}10^{-7}\color{red}units?\color{black}\end{displaymath} 
%\begin{displaymath}{m=}\end{displaymath}  
%\begin{displaymath}{\hbar=}\end{displaymath} 
%\begin{displaymath}{k_B=}\end{displaymath} 
The crystal ideal Helmholtz free energy is computed computationally 
from Equation~\ref{eq:Fideal-n(r)} with the varying number density profile given by
%\begin{align}
  %F_{ideal}[n(\vec{r})]= k_BT\int[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]dV
%\end{align} 
\begin{align}
  {n(|\vec{r}-\vec{R_i}|)=\sum_i{(1-f_v)\frac{e^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{\sigma}^2}}}{\left(\sqrt{2\pi}\sigma\right)^3}}}
\end{align}
%For small gw values it is assumed that the Guassians do not overlap, 
%in which case the equation for the free energy can be simplified, and solved
%analytically with the result
%\begin{equation}
 %F_{Cideal}=(1-f_v)k_BT\left(\ln((1-f_v)\Lambda^3)-3\ln(\sqrt{2\pi{gw}})-\frac{5.0}{2}\right)
%\end{equation}
%For larger gw values it is assumed that the Gaussians do overlap, 
%and so the free energy is caluculated computationally. First dx is scaled by gw,
%\begin{equation}
  %dx_{scaled} = (dx_{input})(gw)
%\end{equation}
%and then $n(\vec{r})$ is computed at one position of $\vec{r}$ within the 
%primitive cell taking into account the contribution of Gaussians from 
%many (5x5x5) nearby cells whose centers are given by $R_i$.
%%\begin{displaymath}{n(|\vec{r}-\vec{R_i}|)+=\frac{(1-f_v)\exp^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}}\end{displaymath}
%The contribution to the free energy of $n(\vec{r})$ at the one position is then
%added to the accumulating sum.
%%\begin{displaymath}{cF_{ideal}+=k_BTn(ln(\frac{(n)2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)dV-DELETE-EQN!}\end{displaymath}
%%\begin{displaymath}{cF_{ideal}+=k_BTn(\ln(n\Lambda^3)-1.0)dV}\end{displaymath}
%This process is repeated for all positions within one primitive cell to get the
%crystal ideal free energy for the primitive cell. This energy is then divided by
%the volume of the primitive cell to get the crystal ideal free energy per
%volume.
%
The liquid excess Helmholtz free energy is computed analytically  
%for a given homogeneous reduced density $n^*$, and reduced temperature $t^*$ 
using SFMT with a uniform number density profile, n that does not depend 
on $\vec r$. 
%done in hf.energy() note- does not depend on fv or gw but does use SFMT):
%\begin{displaymath}{\frac{F_{Hexcess}}{V} = k_BT(\Phi_1+\Phi_2+\Phi_3)}\end{displaymath} 
%For the case of a homogeneous number density n, $\Phi_2$, and $\Phi_3$ simplify to:
%\begin{displaymath}{\Phi_1= -n_0\ln(1-n_3)}\end{displaymath} 
%\begin{displaymath}{\Phi_2= \frac{n_1n_2}{1-n_3}}\end{displaymath} 
%\color{red}\begin{displaymath}{\Phi_3=\frac{n_2^3}{24\pi(1-n_3)^2}}\end{displaymath}\color{black}
%The weighted densities are found from
%\begin{displaymath}{n_i=nw_i}\end{displaymath} 
%where the weights $w_i$ are given by Equations~\ref{eq:weights}.

The crystal excess Helmholtz free energy is computed computationally 
using SFMT with a number density profile $n(\vec r)$ that varies spacially.
%First dx is scaled by $\Xi$ or gw, whichever is larger \begin{displaymath}{dx_{scaled}=(dx_{input})(\Xi + gw)}\end{displaymath}
%Then the crystal, or inhomogeneous, excess free energy is found from
%\begin{displaymath}{F_{Cexcess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{displaymath} 
Monte-Carlo integration is used to compute the weighted densities 
$n_0(\vec{r})$, $n_1(\vec{r})$, $n_2(\vec{r})$, $n_3(\vec{r})$.
%form $\Phi_1(\vec{r})$, $\Phi_2(\vec{r})$, and $\Phi_3(\vec{r})$.
%
The center of with weighting function is placed at some vector $\vec{r}=\vec{r}_k$ 
within the primitive cell, and a vector $\vec{R}_j$ to the center of one of 
the Gaussians is selected that is within the primitive cell, or one of the 
neighboring cells within a prudently picked inclusion radius as Gaussians 
farther away will not contriubute much (see Figure \ref{fig:InclusionRadius}).

\begin{equation}{n_i(\vec{r})= \int_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} 
implemented computationally in the form:
\begin{equation}{n_i(\vec{r})= \sum_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)dV}}\end{equation} 
becomes:
\begin{equation}{n_i(\vec{r})= \sum_{NumPoints}(1-f_v)P_{\vec{dr}}w_i(|\vec{r}-\vec{r}'|)}\end{equation} 
where NumPoints is the number of vectors $\vec{r'}$ randomly generated, and $n_g(\vec{r'})$ 
(the value of the Gaussian at at radius of $|\vec{r'}|$) is equal to a probability density 
which when multiplied by dV gives the probability $P_{\vec{dr}}$ at which the value of 
$w_i(|\vec{r}-\vec{r}'|)$ will occur. 
Using the relation:
\begin{displaymath}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{displaymath} 
$n_i(\vec{r})$ can be expressed as the average value of $(1-f_v)w_i(|\vec{r}-\vec{r}'|)$:
\begin{equation}{n_i(\vec{r})=\sum_{NumPoints}\frac{(1-f_v)w_i(|\vec{r}-\vec{r}'|)}{NumPoints}}\end{equation}
The result can be made more accurate by increasing NumPoints which is the 
number of vectors randomly generated. 
%-------Antithetic-------------------
%In order to further reduce the error 
%in the average, antithetic variates are used. Rather than computing an average 
%of a set of random points, each random point in the set is paired with a 
%point along the same line as the first point, but opposite in direction, 
%as shown in Figure \ref{fig:AntitheticVariate}. This reduces first order 
%error in the average by filtering out an erroneous offset of values along 
%the line connecting the two points by averaging the two results, one of 
%which may be erroneously higher than it should be while the other is 
%erronesously lower than it should be.

 \begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{GaussandW2_thinkas.png}
    \caption{An alternative way of viewing Figure \ref{fig:GaussandW2_actual} where 
    the weight function, shown by the purple dashed lines, is centered at $\vec{r}$  
    and fixed in place as $\vec{dr}$ varies randomly.} 
  \label{fig:GaussandW2_thinkas}
  \end{figure}

@@

%Figure \ref{fig:weight_functions} shows a one dimensional view of the weight functions 
%$w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance along the z-axis. In three 
%dimensions, the weighting functions $w_0$, $w_1$, $w_2$ are spherical shells 
%centered at $\vec{r}$. As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, 
%$w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, a width determined 
%by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving 
%the shell. The weighting function $w_3$, shown in Figure \ref{fig:W0andW3}b is a solid 
%sphere up to $\frac{\alpha-4\Xi}{2}$ after which it tapers off.
%The temperature dependence of parameters $\alpha$ and $\Xi$ 
%%shown in Figures~\ref{fig:alphaXivsT} and \ref{fig:xi_fromB2vsT} 
%can be related 
%to the weight functions. 
%$\Xi$ is related to the width of the functions which 
%become wider in general as the temperature increases. 
%The radial distance 
%to the peak is related to $\alpha$, and decreases as the temperature 
%increases, that is, the shells shrink, as does 
%the sphere.
%The functions can be treated as neglible outside of a distance of 
%about $3(\frac{\Xi}{\sqrt{2}}) + \frac{\alpha}{2}$ CHECK!! from the center of 
%the weight function.

 %\begin{figure}[h!]
    %\centering
    %\includegraphics[height=7cm]{weight_functions.pdf}
    %\caption{The weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
    %z-axis for a reduced temperature of 2.}
    %\label{fig:weight_functions}
  %\end{figure}

 %\begin{figure}[h!]
    %\centering
    %\begin{subfigure}[b]{0.4\linewidth}
      %\includegraphics[height=5cm]{W2.png}
      %% \includegraphics[width=5cm]{figs/w2-ring}
      %\caption{A crossectional view of $w_2$}
    %\end{subfigure} 
    %\begin{subfigure}[b]{0.4\linewidth}
      %\includegraphics[height=5cm]{W3.png}
      %\caption{A crossectional view of $w_3$}
    %\end{subfigure} 
    %\caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
    %spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
    %have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
    %\label{fig:W0andW3}
    %\end{figure}
